Enhancing Underwater Imagery using Generative Adversarial Networks

### 2017-09-16 files/paper/tex/experiments.tex
Final submitted version.

\end{tabular}
\caption{Comparison of local image patches. Each patch was resized to $64 \times 64$.}
\label{fig:zoom}
\end{figure*}

@@

Fig. \ref{mdpmStuff} illustrates the improved performance of MDPM tracker on generated images compared to the real ones. 
Underwater images often fail to capture the true contrast in intensity values between foreground and background due to low 
visibility. The generated images seem to restore these eroded intensity variations to some extent, causing much improved positive 
detection for MDPM tracker.

>>>>

\end{tabular}
\caption{Local image patches extracted for quantitative comparisons, shown in Tables~\ref{fig:gdl_tbl} and~\ref{fig:mean_tbl}. 
Each patch was resized to $64 \times 64$, but shown enlarged for viewing ability.}
\label{fig:zoom}
\end{figure*}

@@

Fig. \ref{mdpmStuff} illustrates the improved performance of MDPM tracker on generated images compared to the real ones. 
Underwater images often fail to capture the true contrast in intensity values between foreground and background due to low 
visibility. The generated images seem to restore these eroded intensity variations to some extent, causing much improved positive 
detection (a 350\% increase in correct detections) for the MDPM tracker.

### 2017-09-16 files/paper/tex/introduction.tex
Final submitted version.

\section{Introduction}
%TODO - talk about having to go back to the same location if you want to get good/bad pairs of the ``same'' image.\starnote{I 
%assume Cam will %do this?}
Underwater robotics has been a steadily growing subfield of autonomous field robotics, assisted by the advent of novel platforms,
sensors and propulsion mechanisms. While autonomous underwater vehicles are often equipped with a variety of sensors, visual 
sensing is an
attractive option because of its non-intrusive, passive, and energy efficient nature. The monitoring of coral reefs 
\cite{shkurti2012multi},
deep ocean exploration \cite{whitcomb2000advances}, and mapping of the seabed~\cite{bingham2010robotic} are a number of tasks 
where
visually-guided AUVs and ROVs (Remotely Operated Vehicles) have seen widespread use. Use of these robots ensures humans are not 
exposed to
the hazards of underwater exploration, as they no longer need to venture to the depths (which was how such tasks were carried out 
in the
past). Despite the advantages of using vision, underwater environments pose unique challenges to visual sensing, as phenomena such 
as light
refraction, absorption and scattering from suspended particles can greatly affect the optics of light. As an example, because red
wavelengths are quickly absorbed by water, images tend to have a green or blue hue to them. As one goes deeper, this effect 
worsens, as more
and more red hue is absorbed. This distortion is extremely non-linear in nature, and is affected by a large number of factors, 
such as the
amount of light present (overcast versus sunny, operational depth), amount of particles in the water, time of day, and the camera 
being
used. This may cause difficulty in tasks such as segmentation, tracking, or classification due to their indirect or direct use of 
color.

As color and illumination begin to change with the depth, vision-based algorithms must need to be very generalizable in order to 
work within the depth ranges a robot may operate in. Because of the high cost and difficulty of acquiring a variety of underwater 
data to train a visual system on, as well as the high amount of noise introduced, algorithms may (and do) perform poorly in these 
different domains. Figure~\ref{fig:samples} shows the high variability in visual scenes that may occur in underwater environments. 
A step towards a solution to this issue is to be able to restore the images such that they appear to be above water, \emph{i.e.}, 
with colors corrected and suspended particles removed from the scene. By performing a many-to-one mapping of these domains from 
underwater to not underwater (what the image would look like above water), algorithms that have difficulty performing across 
multiple forms of noise may be able to focus only one clean domain.

@@

   \includegraphics[width=0.8in]{n01496331_7428_f1} &
   \includegraphics[width=0.8in]{n01496331_16340_f1} &
   \includegraphics[width=0.8in]{n01914609_5148_f1} &
   \includegraphics[width=0.8in]{robot_f1} \\
   \includegraphics[width=0.8in]{n01664065_30279_f1} &
   \includegraphics[width=0.8in]{n01917289_5711_f1} &
   \includegraphics[width=0.8in]{n01917289_4087_f1} &
   \includegraphics[width=0.8in]{n02607072_10395_f1} \\
\end{tabular}
\label{fig:samples}
\caption{Sample underwater images with natural and man-made artifacts (which in this case is our underwater robot) displaying the 
diversity of distortion that can occur. With the varying camera-to-object distances in the images, the distortion and loss of 
color varies between the different images.}
\vspace{-5mm}
\end{figure}

>>>>

\section{Introduction}
%TODO - talk about having to go back to the same location if you want to get good/bad pairs of the ``same'' image.\starnote{I 
%assume Cam will %do this?}
Underwater robotics has been a steadily growing subfield of autonomous field robotics, assisted by the advent of novel platforms, 
sensors and propulsion mechanisms. While autonomous underwater vehicles are often equipped with a variety of sensors, visual 
sensing is an attractive option because of its non-intrusive, passive, and energy efficient nature. The monitoring of coral reefs 
\cite{shkurti2012multi}, deep ocean exploration \cite{whitcomb2000advances}, and mapping of the seabed~\cite{bingham2010robotic} 
are a number of tasks where visually-guided AUVs and ROVs (Remotely Operated Vehicles) have seen widespread use. Use of these 
robots ensures humans are not exposed to the hazards of underwater exploration, as they no longer need to venture to the depths 
(which was how such tasks were carried out in the past). Despite the advantages of using vision, underwater environments pose 
unique challenges to visual sensing, as light refraction, absorption and scattering from suspended particles can greatly affect 
optics. For example, because red wavelengths are quickly absorbed by water, images tend to have a green or blue hue to them. As 
one goes deeper, this effect worsens, as more and more red hue is absorbed. This distortion is extremely non-linear in nature, and 
is affected by a large number of factors, such as the amount of light present (overcast versus sunny, operational depth), amount 
of particles in the water, time of day, and the camera being used. This may cause difficulty in tasks such as segmentation, 
tracking, or classification due to their indirect or direct use of color.

As color and illumination begin to change with depth, vision-based algorithms need to be generalizable in order to work within 
the depth ranges a robot may operate in. Because of the high cost and difficulty of acquiring a variety of underwater data to 
train a visual system on, as well as the high amount of noise introduced, algorithms may (and do) perform poorly in these 
different domains. Figure~\ref{fig:samples} shows the high variability in visual scenes that may occur in underwater environments. 
A step towards a solution to this issue is to be able to restore the images such that they appear to be above water, \emph{i.e.}, 
with colors corrected and suspended particles removed from the scene. By performing a many-to-one mapping of these domains from 
underwater to not underwater (what the image would look like above water), algorithms that have difficulty performing across 
multiple forms of noise may be able to focus only one clean domain.

@@

   \includegraphics[width=0.8in]{n01496331_7428_f1} &
   \includegraphics[width=0.8in]{n01496331_16340_f1} &
   \includegraphics[width=0.8in]{n01914609_5148_f1} &
   \includegraphics[width=0.8in]{robot_f1} \\
   \includegraphics[width=0.8in]{n01664065_30279_f1} &
   \includegraphics[width=0.8in]{n01917289_5711_f1} &
   \includegraphics[width=0.8in]{n01917289_4087_f1} &
   \includegraphics[width=0.8in]{n02607072_10395_f1} \\
\end{tabular}
\label{fig:samples}
\caption{Sample underwater images with natural and man-made artifacts (which in this case is our underwater robot) displaying the 
diversity of distortions that can occur. With the varying camera-to-object distances in the images, the distortion and loss of 
color varies between the different images.}
\vspace{-5mm}
\end{figure}

### 2017-09-16 files/paper/tex/methodology.tex
Final submitted version.

\section{Methodology}
\label{sec:methodology}
Underwater images distorted by color or circumstance lack ground truth, which is a necessity for previous colorization
approaches. Furthermore, the distortion present in an underwater image is highly nonlinear; simple methods such as
adding a hue to an image do not capture all of the dependencies. We propose to use CycleGAN as a distortion model in
order to generate paired images for training. Given a domain of underwater images with no distortion, and a domain of
underwater images with distortion, CycleGAN is able to perform style transfer. Given an undistorted image, CycleGAN
distorts it such that it appears to have come from the domain of distorted images. These pairs are then used in our
algorithm for image reconstruction.

@@

To circumvent the problem of insufficient image pairs, we use CycleGAN to generate $I^D$ from $I^C$, which gives us a 
paired dataset of images. Given two datasets $X$ and $Y$, where $I^C \in X$ and $I^D \in Y$, CycleGAN learns a mapping $F: X 
\rightarrow Y$. Figure~\ref{fig:cgan_samples} shows paired samples generated from CycleGAN. From this paired dataset we train a 
generator $G$ to learn the function $f: I^D \rightarrow I^C$. It should be noted that during our data generation process, CycleGAN 
simultaneously learns a mapping $G: Y \rightarrow X$, which is similar to $f$. In Section~\ref{sec:experiments}, we compare images 
generated by CycleGAN with images generated through our approach.

\subsection{Adversarial Networks}
\label{sec:gans}
\nostarnote{check this paragraph as I added some more text}In machine learning literature, Generative Adversarial Networks 
(GANs)~\cite{goodfellow2014generative} represent a class of generative models based on game theory in which a generator network 
competes against an adversary. From a classification perspective, the generator network produces instances which actively attempt 
to `fool' the discriminator network. The goal is for the discriminator network to be able to distinguish between `true' 
instances coming from the dataset and `false' instances produced by the generator network.
%, and thus improve its classification accuracy.
In our case, conditioned on an image $I^D$, the generator is trained to produce an image to try and fool the 
discriminator, which is trained to distinguish between distorted and non-distorted underwater images. In the original
GAN formulation, our goal is to solve the minimax problem:

@@

Note for simplicity in notation, we will further omit $I^C \sim p_{train}$ and $I^D \sim p_{gen}$. In this formulation, 
the discriminator is hypothesized as a classifier with a sigmoid cross-entropy loss function, which in practice may lead
to issues  such as the vanish gradient and mode collapse. As shown by \cite{arjovsky2017towards}, as
the discriminator improves, the gradient of the generator vanishes, making it difficult or impossible to train. Mode
collapse occurs when the generator ``collapses'' onto a single point, fooling the discriminator with only one instance.
To illustrate the effect of mode collapse, imagine a GAN is being used to generate digits from the MNIST~\cite{lecun2010mnist} 
dataset, but it only generated the same digit. In reality, the desired outcome would be to generate a diverse collection of all 
the digits.
There have been a number of recent methods which hypothesize a different loss function for the discriminator
\cite{mao2016least,arjovsky2017wasserstein,gulrajani2017improved,zhao2016energy}. We focus on the Wasserstein GAN
(WGAN) \cite{arjovsky2017wasserstein} formulation, which proposes to use the Earth-Mover or \textit{Wasserstein-1}
distance $W$ by constructing a value function using the Kantorovich-Rubinstein duality \cite{villani2008optimal}.
In this formulation, $W$ is approximated given a set of $k$-Lipschitz functions $f$ modeled as neural networks. To
ensure $f$ is $k$-Lipschitz, the weights of the discriminator are clipped to some range $[-c, c]$. In our work, we
adopt the Wasserstein GAN with gradient penalty (WGAN-GP) \cite{gulrajani2017improved}, which instead of clipping
network weights like in \cite{arjovsky2017wasserstein}, ensures the Lipschitz constraint by enforcing a soft
constraint on the gradient norm of the discriminator's output with respect to its input. Following
\cite{gulrajani2017improved}, our new objective then becomes

@@

\subsection{Image Gradient Difference Loss}
Often times generative models produce blurry images. We explore a strategy to sharpen these predictions by
directly penalizing the differences of image gradient predictions in the generator, as proposed by
\cite{mathieu2015deep}. Given a ground truth image $I^C$, predicted image $I^P = G(I^D)$, and $\alpha \geq 1$, the
Gradient Difference Loss (GDL) is given by

>>>>

\section{Methodology}
\label{sec:methodology}
Underwater images distorted by color or other circumstances lack ground truth, which is a necessity for previous colorization
approaches. Furthermore, the distortion present in an underwater image is highly nonlinear; simple methods such as
adding a hue to an image do not capture all of the dependencies. We propose to use CycleGAN as a distortion model in
order to generate paired images for training. Given a domain of underwater images with no distortion, and a domain of
underwater images with distortion, CycleGAN is able to perform style transfer. Given an undistorted image, CycleGAN
distorts it such that it appears to have come from the domain of distorted images. These pairs are then used in our
algorithm for image reconstruction.

@@

To circumvent the problem of insufficient image pairs, we use CycleGAN to generate $I^D$ from $I^C$, which gives us a 
paired dataset of images. Given two datasets $X$ and $Y$, where $I^C \in X$ and $I^D \in Y$, CycleGAN learns a mapping $F: X 
\rightarrow Y$. Figure~\ref{fig:cgan_samples} shows paired samples generated from CycleGAN. From this paired dataset we train a 
generator $G$ to learn the function $f: I^D \rightarrow I^C$. It should be noted that during the training process of CycleGAN, 
it simultaneously learns a mapping $G: Y \rightarrow X$, which is similar to $f$. In Section~\ref{sec:experiments}, we compare 
images generated by CycleGAN with images generated through our approach.

\subsection{Adversarial Networks}
\label{sec:gans}
\nostarnote{check this paragraph as I added some more text}In machine learning literature, Generative Adversarial Networks 
(GANs)~\cite{goodfellow2014generative} represent a class of generative models based on game theory in which a generator network 
competes against an adversary. From a classification perspective, the generator network $G$ produces instances which actively 
attempt 
to `fool' the discriminator network $D$. The goal is for the discriminator network to be able to distinguish between `true' 
instances coming from the dataset and `false' instances produced by the generator network.
%, and thus improve its classification accuracy.
In our case, conditioned on an image $I^D$, the generator is trained to produce an image to try and fool the 
discriminator, which is trained to distinguish between distorted and non-distorted underwater images. In the original
GAN formulation, our goal is to solve the minimax problem:

@@

Note for simplicity in notation, we will further omit $I^C \sim p_{train}(I^C)$ and $I^D \sim p_{gen}(I^D)$. In this formulation, 
the discriminator is hypothesized as a classifier with a sigmoid cross-entropy loss function, which in practice may lead
to issues  such as the vanish gradient and mode collapse. As shown by \cite{arjovsky2017towards}, as
the discriminator improves, the gradient of the generator vanishes, making it difficult or impossible to train. Mode
collapse occurs when the generator ``collapses'' onto a single point, fooling the discriminator with only one instance.
To illustrate the effect of mode collapse, imagine a GAN is being used to generate digits from the MNIST~\cite{lecun2010mnist} 
dataset, but it only generated the same digit. In reality, the desired outcome would be to generate a diverse collection of all 
the digits. To this end, there have been a number of recent methods which hypothesize a different loss function for the 
discriminator
\cite{mao2016least,arjovsky2017wasserstein,gulrajani2017improved,zhao2016energy}. We focus on the Wasserstein GAN
(WGAN) \cite{arjovsky2017wasserstein} formulation, which proposes to use the Earth-Mover or \textit{Wasserstein-1}
distance $W$ by constructing a value function using the Kantorovich-Rubinstein duality \cite{villani2008optimal}.
In this formulation, $W$ is approximated given a set of $k$-Lipschitz functions $f$ modeled as neural networks. To
ensure $f$ is $k$-Lipschitz, the weights of the discriminator are clipped to some range $[-c, c]$. In our work, we
adopt the Wasserstein GAN with gradient penalty (WGAN-GP) \cite{gulrajani2017improved}, which instead of clipping
network weights like in \cite{arjovsky2017wasserstein}, ensures the Lipschitz constraint by enforcing a soft
constraint on the gradient norm of the discriminator's output with respect to its input. Following
\cite{gulrajani2017improved}, our new objective then becomes

@@

\subsection{Image Gradient Difference Loss}
Often times generative models produce blurry images. We explore a strategy to sharpen these predictions by
directly penalizing the differences of image gradient predictions in the generator, as proposed by
\cite{mathieu2015deep}. Given a ground truth image $I^C$, predicted image $I^P = G(I^D)$, and $\alpha$ which is an integer 
greater than or equal to 1, the Gradient Difference Loss (GDL) is given by
